---
title: "Mitchell River"
output:
  rmarkdown::html_document:
    self_contained: yes
fontsize: 11pt
documentclass: article
vignette: >
  %\VignetteIndexEntry{Mitchell River}
  %\VignetteEngine{knitr::rmarkdown_notangle}
---

```{r setup, include = FALSE}
h <- 6
w <- 8
is_check <- ("CheckExEnv" %in% search()) || any(c("_R_CHECK_TIMINGS_",
             "_R_CHECK_LICENSE_") %in% names(Sys.getenv()))
knitr::opts_chunk$set(
  fig.align = "center", 
  eval = !is_check,
  collapse = TRUE,
  comment = "#>")
```

The Mitchell River is a river located in Far North Queensland, Australia. The river rises on the Atherton Tableland about 50 kilometres (31 mi) northwest of Cairns, and flows about 750 kilometres (470 mi) northwest across Cape York Peninsula from Mareeba to the Gulf of Carpentaria. we will use this case study to present some functionalities of the ´prioriactions´ package.

We started loading libraries:

```{r eval=FALSE, message=FALSE, warning=FALSE}
library(prioritizr) #To use the simulate features distribution
library(prioriactions)
library(raster) #To plot of shapefiles
library(tmap) #To create cool maps
library(scales) #To standardize the value of amount
library(reshape2) #To use the melt function
library(sp) #To use the spplot function
library(viridis) #To use viridis pallete
```

# Preparing data inputs

We divided the whole catchment (71,630 km2 into 2316 sites (i.e., sub-catchments), each one included the portion of river length between two consecutive river connections. We considered four major threats to freshwater fish species in the catchment: water buffalo (Bubalis bubalis), cane toad (Bufo marinus), river flow alteration (caused by impoundments, channels for water extractions and levee banks) and grazing land use. Also, we used the modelled spatial distribution of 45 fish species in the Mitchell river catchment as our conservation features. The distribution of features and threats () will be simulated while the other input files a *prioriactions* will be loaded directly from the package as follows:
